Simple and robust inference of connectomes using partial correlation coefficients

### 2014-06-09 article.tex
Some modifs to the proposition of alternate version + Alternate version of inference process. Note: Both versions should be merged.

\textcolor{red}{AS: Après avoir discuté avec Gilles, une autre façon de présenter les choses qui est moins rigoureuse que l'approche de AJ mais plus "motivée"}
Prior inferring the network, the fluorescent calcium time series are smoothed
using several linear and non linear filters.

@@

We thus define our preprocessing function $g$ as 
\begin{align}
g(X_t^i) = g_1(g_3(g_2(X_t^i))).
\end{align}

@@

\hrule

@@

\textcolor{red}{Motiviation pour l'inverse : retirer les liens indirects} \\

>>>>

\textcolor{red}{AS: Après avoir discuté avec Gilles, une autre façon de
présenter les choses qui est moins rigoureuse que l'approche de AJ mais plus
"motivée" --> Certains aspects pourraient être incorporés dans la version de AJ.}

Prior inferring the network, the fluorescent calcium time series
are smoothed using several linear and non linear filters.

@@

@@

Hence, we define our preprocessing function $G$ as 
\begin{align}
G(X_t^i) = g_5 \bullet g_4 \bullet g_1 \bullet g_3 \bullet g_2(X_t^i).
\end{align}

\vspace{1em}
\hrule
\vspace{1em}

@@

\textcolor{red}{Motivation pour l'inverse : retirer les liens indirects. Pareil: proposition de AS  et certains choses pourraient être incorporées dans la version de AJ.} \\

From a probabilistic point of view, the inference of the undirected network
can be formulated as inferring conditional dependences and independences. Even
if \textit{correlation does no imply causation}[todo cite], it gives insight
on the underlying connectivity network.

Let us consider see neurons as random variables $\mathcal{X}_i$ which follows
a probability density function $p_{\mathcal{X}_i}$ and let us denote by
$p_{\mathcal{X}_i,\mathcal{X}_j|\mathcal{Z}}$ the joint probability
distribution between neurons $i$ and $j$ conditioned on a subset of neurons
$\mathcal{Z}$ (eventually empty). Let us define $\mathcal{V}$ the set of all
neurons $\mathcal{X}_1, \dots, \mathcal{X}_p$, and by $\mathcal{V}^{i,j}$ the
same set without neurons $\mathcal{X}_i$ and $\mathcal{X}_j$.

 First, if we want to highlight
relationships between without considering the distinction between direct and
indirect links, the Pearson correlation matrix, i.e. each element $(i,j)$ of a
such matrix is the correlation coefficient between neurons $\mathcal{X}_i$ and $\mathcal{X}_j$.

However the connectivity network to infer is only made of direct relationships
which can be defined as a relation between neurons that still exists when
conditioned on all other neurons or, in other words, there is not intermediary
between those two neurons. 
Under the hypothesis that neurons follow a jointly Gaussian distribution $p_{\mathcal{X}_1, \dots, \mathcal{X}_p} \sim \mathcal{N}(\mu; \Sigma)$, two neurons $\mathcal{X}_i$ and $\mathcal{X}_j$ are independent $\mathcal{X}_i\perp \mathcal{X}_j$ if the Pearson correlation coefficient $\rho_{X_iX_j} = \frac{\Sigma_{ij}}{\sqrt{\Sigma_{ii} \Sigma_{jj}}} = 0,$
and are conditionally independant to all others neurons $\mathcal{X}_i \perp \mathcal{X}_j |\mathcal{V}^{i,j}$ if the partial correlation coefficient (of order $p-2$) $\rho_{X_i, X_j | \mathcal{V}^{i,j}} =
\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.$ 
Therefore, by definition, the partial correlation coefficient between two neurons only estimates the direct relationship between them.

In previous study \cite{shipley2002cause}, it has been suggested to
consider correlation coefficients of all possible orders,
i.e. conditioned on every subsets of the set of $p-2$ other variables. This
would take into account multiple indirect paths between a given pair of
variables. An obvious limitation to estimate correctly high-order coefficients
is the number of samples. However given the high number of samples $n$
in the Connectomics challenge, we only consider the highest order of partial
correlation coefficient.

\vspace{1em}
\hrule
\vspace{1em}
